Enzo: An Adaptive Mesh Refinement Code for Astrophysics

### 2013-06-20 parallel.tex
updated GPU section a bit

To illustrate the speedup provided by porting two solvers to GPUs, we
show the results of a weak scaling test of driven turbulence in
Figure~\ref{fig:gpu_scaling}.  This \enzo\ problem type uses the
Dedner MHD solver, which runs on both CPUs and GPUs, and has been
tuned to maximize performance on both platforms.  The benchmarking platform is a Cray
XK6 system where each node has a 16-core AMD Opteron 6272 CPU and a single
NVIDIA Tesla K20 GPU, which has 2,496 CUDA cores and a maximum
theoretical speed of 1.17 Tflops when used in double precision.
For the CPU run, all the 16 cores in each node
are used by launching 16 MPI processes per node. For the GPU run, 16 MPI
processes are also launched per node, and all processes on this node share the single GPU
on that node. This can work because NVIDIA's MPS (Multi-Process
Service) technology allows multiple processes to concurrently use the
same GPU.  A series of weak scaling tests were run, varying from 1
to 8 nodes (16 to 128 MPI processes), with the number of cells per
node fixed to $256^3$.  Figure~\ref{fig:gpu_scaling} displays the
scaling results in terms of cell updates per second per node -- as a
result, ideal weak scaling is a horizontal line.  In general, the
simulations using the GPU-accelerated MHD solver perform roughly 5
times better overall on this system than the CPU-only calculations.

>>>>

To illustrate the speedup provided by porting two solvers to GPUs, we
show the results of a weak scaling test of driven MHD turbulence on a
uniform mesh in Figure~\ref{fig:gpu_scaling}.  This particular problem
type contains no physics other than solving the equations of ideal
MHD, and thus is representative of the type of calculation than can
benefit from the GPU-optimized solvers in \enzo.  In this scaling
test, we use the Dedner MHD solver, which runs on both CPUs and GPUs
and has been tuned to maximize performance on both platforms.  The
benchmarking platform is a Cray XK6 system with a Gemini
interconnect. Each node has a 16-core AMD Opteron 6272 CPU and a
single NVIDIA Tesla K20 GPU, which has 2,496 CUDA cores and a maximum
theoretical speed of 1.17 Tflops when used in double precision.  For
the CPU run, all the 16 cores in each node are used by launching 16
MPI processes per node. For the GPU run, 16 MPI processes are also
launched per node, and all processes on this node share the single GPU
on that node. This can work because NVIDIA's MPS (Multi-Process
Service) technology allows multiple processes to concurrently use the
same GPU.  A series of weak scaling tests were run, varying from 1 to
8 nodes (16 to 128 MPI processes), with each node containing a cube of
$256^3$ cells (so, a 4-node computation would have a domain that is
$512 \times 512 \times 256$ cells).  Figure~\ref{fig:gpu_scaling}
displays the scaling results in terms of cell updates per second per
node -- as a result, ideal weak scaling is a horizontal line.  In
general, the simulations using the GPU-accelerated MHD solver perform
roughly 5 times better overall on this system than the equivalent
CPU-only calculation run on the same number of nodes.  We note,
however, that this somewhat mis-represents the speedup obtained by
porting the MHD code to GPUs -- in the GPU simulation case, the 16 CPU
cores on each node are effectively idle, as they are used only for
timestep calculation and boundary condition transmission (which are
both computationally inexpensive compared to the fluid solve).  This
means that, effectively, the GPU simulation is 80 times faster than a
single CPU core, so a user would likely see much greater effective
speedup on a system where the ratio of CPU cores to GPUs is lower.
